---
title: "Take-home_Ex02"
author: "WANG XINGYUN"
---

# **Take-home Exercise 2: Be Tradewise or Otherwise**

## **Overview**

We try to apply newly acquired techniques to explore and analyze the changing trends and patterns of Singapore’s international trade since 2015.

### **Setting the scene**

## Our task

## Getting started

### load package

For this exercise, we load the following R packages using the `pacman::p_load()` function, R packages including: scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table and tidyverse.

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               CGPfunctions, ggHoriPlot, tidyverse)
```

we can download 3 files (M451491, M451501, M451511) from the website (download [Merchandise Trade by Region/Market](https://www.singstat.gov.sg/find-data/search-by-theme/trade-and-investment/merchandise-trade/latest-data) from [Department of Statistics Singapore, DOS](https://www.singstat.gov.sg/)), and the 3 files represent different data, as shown below

M451491-Merchandise Trade By Region and Selected Market (Imports), Monthly

M451501-Merchandise Trade By Region and Selected Market (Domestic Exports),Monthly

M451511-Merchandise Trade By Region and Selected Market (Re-Exports), Monthly

### Import data

```{r}
df_Im <- read_csv("data/M451491.csv", skip = 10, show_col_types = FALSE)
head(df_Im)
```

```{r}
df_DE <- read_csv("data/M451501.csv", skip = 10, show_col_types = FALSE)
df_RE <- read_csv("data/M451511.csv", skip = 10, show_col_types = FALSE)
head(df_DE)
head(df_RE)
```

After viewing the whole table, we should change the type of date series.

First column (Data Series): Country/Region Name, Subsequent columns: Monthly data, such as "2025 Jan", "2024 Dec", "2024 Nov", representing trade data for different months.

The column name is in YYYY Mon format (needs to be converted to standard time format YYYY-MM). Data needs to be converted to Long Format for time series analysis.

```{r}
# rename "Data Series" as "Country"
colnames(df_Im)[1] <- "Country"
colnames(df_DE)[1] <- "Country"
colnames(df_RE)[1] <- "Country"

# Extract time column names and convert them to YYYY-MM format
time_cols <- colnames(df_Im)[-1]  
time_cols_cleaned <- format(parse_date_time(time_cols, orders = "ym"), "%Y-%m")

colnames(df_Im)[-1] <- time_cols_cleaned
colnames(df_DE)[-1] <- time_cols_cleaned
colnames(df_RE)[-1] <- time_cols_cleaned

head(df_Im)
```

At present, the data is in Wide Format, and we need to convert it to Long Format for time series analysis:

```{r}
# 转换为长表格式
df_Im_long <- df_Im %>%
  pivot_longer(cols = -Country, names_to = "Date", values_to = "Trade_Value")

df_DE_long <- df_DE %>%
  pivot_longer(cols = -Country, names_to = "Date", values_to = "Trade_Value")

df_RE_long <- df_RE %>%
  pivot_longer(cols = -Country, names_to = "Date", values_to = "Trade_Value")

# 确保 Date 列是日期格式
df_Im_long$Date <- as.Date(paste0(df_Im_long$Date, "-01"))  # 转换为 YYYY-MM-DD
df_DE_long$Date <- as.Date(paste0(df_DE_long$Date, "-01"))
df_RE_long$Date <- as.Date(paste0(df_RE_long$Date, "-01"))

# 查看转换后的数据
head(df_Im_long)

```

Save the cleaned data

```{r}
write_csv(df_Im_long, "data/cleaned_Im.csv")
write_csv(df_DE_long, "data/cleaned_DE.csv")
write_csv(df_RE_long, "data/cleaned_RE.csv")
```
